---
title: 'Homework #2'
author: "Guang Yi Lim; gylim0604"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This homework is about syntax and basic plotting functions in R, and is **due on Feb 1** before 11:59pm.

You can use the RMarkDown file with these homework instructions as a template for your work. 
Make sure to replace "Your Name" and "your_github_username" in the YAML with your name and Github username.

**Submission instructions**: Create a folder named `ds202_hw2`, and name the RMarkDown file including your solutions `hw2.Rmd` under this folder. For submission, create a GitHub repository named `ds202_hw2` under your account, and push both `hw2.Rmd` and the knitted `hw2.html` to GitHub before the deadline. I will assume you use the same GitHub username as for your HW1 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. Some questions will require the following answer format.

## Example question and answer
0\. What is the smallest positive odd integer? Storage it in a variable named `a0` and print it.

**Example answer**:
In the RMarkDown file, type in
<pre><code>```{r}
a0 <- 1
a0
```</code></pre>
The knitted html file will show:
```{r}
a0 <- 1
a0
```

## R commands
1. Today's temperature has a high of 15 and a low of -10 degree fahrenheit.  Find the corresponding degress in celsius and storage it in a (single) vector named `cel`. Print the variable.
```{r}
  cel <- (1:2)
  cel[1] = (15 -32) *5/9
  cel[2] = (-10-32) * 5/9
  print(cel)
```
2. Create a vector of length 200. The odd entries should be 1 and even entries should be 0. Storage the created in a variable named `vec200` and print.
```{r}
  vec200 <-(1:200)
  for(i in 1:200){
    if(i%%2!=0){
      vec200[i] = 1;
    }
    else{
      vec200[i]=0;
    }
  }
  print(vec200);
```
3. How many perfect squares are there between 1 and 1001? Storage the answer in a  variable named `numPerfect` and print. You may want to look at `?round`.
```{r}
#need help pls send nudes
  numPerfect <- 0;
  for(i in 1:1001){
    for(j in 1:i){
      if(j*j ==i){
        numPerfect <- numPerfect+1;
      }
    }
  }
  print(numPerfect)
```
## Cars and mileage
For this part we will use the `mpg` data set, you can access the data by loading the package `ggplot2` into your R session by executing the command 
`library(ggplot2)`. The `mpg` data includes information about fuel economy of popular car models
for the years 1999 and 2008 as collected by the US Environmental Protection Agency. An answer should include a short paragraph of text for the explanation to a (human) reader, together with appropriate R code, outputs, and graphs.

Most exercises are from the ggplot2 book section 2.2.1. You might find it helpful for working on the homework to read through this section.
```{r}
  library(ggplot2)
```
1. What are the 3 most fuel-efficient models in terms of highway mileage? Storage the data frame (with 3 rows) containing all information of these three cars in a variable named `top3` and print.
```{r}
  temp <-mpg[order(mpg$hwy, decreasing = TRUE),]
  top3 <- temp[1:3,]
  print(top3)
```
2. How many different compact models are included in this dataset? Storage the answer in  a variable named `numCompact` and print.
```{r}
  temp <-subset(mpg, mpg$class == 'compact')
  numCompact = nrow(temp)
  print(numCompact)
```
3. Create a scatterplot of `hwy` and `cty` and describe the relationship. Why are there so few points visible?
```{r}
  plot(mpg$hwy, mpg$cty,pch = c(1,2),col = c("red","blue"))
  legend("topleft",legend = c("hwy","cty"),pch = c(1,2),col = c("red","blue"))
```
</br>
4. Are cars in 2008 more fuel efficient than cars in 1999? Use graphs *and* summary statistics to answer this question. 
```{r}
  d2008 <-subset(mpg,mpg$year ==2008)
  d1999 <-subset(mpg,mpg$year ==1999)
  scatter <- c(d2008,d1999)
  plot(d2008$hwy,d1999$hwy,pch =c(16,17),col =c("red","blue"),main="Highway miles per gallon",xlab = "2008",ylab = "1999")
  legend("topleft",legend = c("2008","1999"),col = c("red","blue"),pch = c(16,17))
  plot(d2008$cty,d1999$cty,pch =c(16,17),col =c("red","blue"),main = "City miles per gallon",xlab = "2008",ylab ="1999")
  legend("topleft",legend = c("2008","1999"),col = c("red","blue"),pch = c(16,17))
  boxplot(c(d2008$hwy, d2008$cty),c(d1999$hwy,d1999$hwy),main ="Car Milage Data",xlab = "Year", ylab = "Miles per gallon",col = c("red","blue")) 
  legend("topleft",legend = c("2008","1999"),col = c("red","blue"),pch= 15)
#looking at the the two scatterplots, we can see that cars in 1999 are more fuel efficient than cars in 2008 in the city miles per gallon. This is also backed by the  boxplot of cars milage data.
  
```
</br>
Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission. Make sure that you don't use any file structure that depends on your computing environment.